---
title: "DATA607_Project3_Wrangling"
author: "Josh Iden"
date: "2022-10-20"
output:
  pdf_document: default
  html_document:
    includes:
      in_header: header.html
    css: ./lab.css
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DATA WRANGLING

```{r, warning=FALSE, message=FALSE}
# load packages
library(tidyverse)
library(readxl)
library(stringr)
library(lubridate)
```

We stored the original multi-sheet .xlsx file from Kaggle in the project GitHub repo. Due to the structure of the file, we extracted the individual sheets locally, which we then wrote to CSVs and stored to the project repo for database ingestion:

```{r load-data}
# read file locally 
file <- "/Users/joshiden/Documents/Classes/CUNY SPS/Fall 2022/DATA 607/Projects/Project 3/Data Science Career Terms.xlsx"
excel <- read_excel(file)

# store sheet names
sheets <- excel_sheets(file)

# read sheets into dataframes
ds_skills <- read_excel(file, sheet = sheets[1])
ds_software <- read_excel(file, sheet = sheets[2])
education <- read_excel(file, sheet=sheets[7])
```

Previewing the output dataframes,

```{r}
head(ds_skills)
```

```{r}
head(ds_software)
```

```{r}
head(education)
```

Tidy technical and soft skills for category output tables. **Keyword** column contains skill names from first row until cell value **Total**.

```{r tidy-skills}
# Find row index of Total
totalIdx <- which(ds_skills$Keyword == "Total")

skills <- ds_skills |> 
           # Grab beginning rows until and excluding Total entry
           slice_head(n = totalIdx - 1) |>
           # select Keyword column
           select(Keyword) |>
           # drop NA
           filter(! is.na(Keyword)) |>
           # uppercase
           mutate(Keyword = str_to_upper(Keyword)) |>
           # add T_GENERAL and T_SOFT categories
           mutate(Category=ifelse(Keyword == "COMMUNICATION", "T_SOFT", 
                    ifelse(Keyword == "PROJECT MANAGEMENT", "T_SOFT", "T_GENERAL")))
nrow(skills)
tail(skills)
```

```{r tidy-software}
totalIdx <- which(ds_software$Keyword == "Total")

software <- ds_software |> 
           # Grab beginning rows until and excluding Total entry
           slice_head(n = totalIdx - 1) |> 
           # select Keyword column
           select(Keyword) |>
           # drop NA
           filter(! is.na(Keyword)) |>
           # uppercase
           mutate(Keyword = str_to_upper(Keyword)) |>
           # add T_SOFTWARE category
           mutate(Category = "T_SOFTWARE")
nrow(software)
tail(software)
```

Applying transformations to standardize and retain the desired data for data tables:

```{r transform-skills}
# keep only first 15 rows
# Keyword to upper
# pivot columns to column: source
ds_skills_transformed <- ds_skills |>
  head(15) |>
  mutate(Keyword = toupper(Keyword)) |>
  rename(KEYWORD = Keyword) |>
  pivot_longer(cols=("LinkedIn":"Monster"), names_to="SOURCE", values_to="COUNT") |>
  mutate(SOURCE = toupper(SOURCE), SURVEY_DATE=ymd("2018-06-15")) |>
  arrange(KEYWORD,SOURCE)

ds_skills_transformed
```

```{r transform-software}
# keep top 37 rows
# keyword to upper
# pivot columns to source
# source column to upper
# add date column
ds_software_transformed <- ds_software |>
  select(c("Keyword":"Monster")) |>
  head(37) |>
  mutate(Keyword = toupper(Keyword)) |>
  rename(KEYWORD = Keyword) |>
  pivot_longer(cols=("LinkedIn":"Monster"), names_to="SOURCE", values_to="COUNT") |>
  mutate(SOURCE = toupper(SOURCE), SURVEY_DATE=ymd("2018-06-15")) |>
  arrange(KEYWORD,SOURCE)

ds_software_transformed
```

```{r transform-education}
# keyword to uppercase
# pivot columns to source
# source column to uppercase
# add date column
# drop AngelList column
# drop NA values
education_transformed <- education |>
  mutate(Keyword = toupper(Keyword)) |>
  rename(KEYWORD = Keyword) |>
  pivot_longer(cols=("LinkedIn":"SimplyHired"), names_to="SOURCE", values_to="COUNT") |>
  mutate(SOURCE = toupper(SOURCE), SURVEY_DATE=ymd("2018-06-15")) |>
  subset(select = -c(AngelList)) |>
  drop_na() |>
  arrange(KEYWORD,SOURCE)

education_transformed
```

Combining ds_skills_transformed and ds_software_transformed -\> skills_in_demand

```{r transform-skills-in-demand}
#skills in demand
skills_in_demand <- rbind(ds_skills_transformed,ds_software_transformed)
skills_in_demand
```

The files were then written to CSV and added to the project repository for database ingestion.

```{r output-files}
# write to software_skills.csv
# outputDir <- "/Users/jawaidhakim/Downloads/"
# outputFile <- str_c(inputDir, "software_skills.csv")
# write.csv(software, outputFile)

# write to general_skills.csv
# outputFile <- str_c(inputDir, "general_skills.csv")
# write.csv(skills, outputFile)

# write to skills_in_demand.csv
# sid <- "/Users/joshiden/Documents/Classes/CUNY SPS/Fall 2022/DATA 607/Projects/skills_in_demand.csv"
# write.csv(skills_in_demand, sid)

# write to education_in_demand.csv
# eid <- "/Users/Melissa/OneDrive/Documents/CUNY/Fall 2022/Data 607/Project 3/education_in_demand.csv"
# write.csv(education_transformed, eid)
```
